Weather_covariance_matrices_flowering

### 2020-12-03 R/Analysis_v0.5_weather_covariance_Uhyp_flowering.Rmd
Add analysis with weather phenotypes as covariance U_hyp

  ggcov1 <- pvdiv_plot_cov(cor_phe) + ggtitle(label = paste0(phe_hyp[i], "_", names(subpop_v2)[k]))
  ggcov <- list.prepend(ggcov, ggcov1)
  U_hyp <- list.prepend(U_hyp, cor_phe)
  names(U_hyp)[1] <- paste0(phe_hyp[i], "_", names(subpop_v2)[k])
  }
}

@@

## Midwest NB
Some univariate GWAS for the Midwest had population structure issues. Mostly for greenup - even GR50 at MO was quite bad. Drop MO for GR50. Otherwise they look ok for FL50 and GR50.

## Set up mash inputs
```{r}
i=1 # site
j=1 # phe
k=3 # subpop: G&M, M, Gulf
for(k in c(1:3)){
phe_suffix <- paste("FL50", subpop_v[k], 
                    sep = "*.*")
snp <- snp_attach(inputfiles$SNPfiles[k])
gwas_rds <- pvdiv_results_in_folder(path = file.path(outputdir), 
                                      pattern = paste0("GWAS_datatable*.*",
                                                       phe_suffix))
phenotypes_1 <- str_sub(gwas_rds, start = 16, end = -42) %>%
    str_replace(., "_$", "")
phenotypes <- paste0(names(subpop_v)[k], "_", phenotypes_1)
numSNPs <- round(ceiling(1200000/length(phenotypes)^2)/1000)*1000
out1phe <- pvdiv_bigsnp2mashr(path = file.path(outputdir), 
                           snp = snp, gwas_rds = gwas_rds, 
                           phenotypes = phenotypes, numSNPs = numSNPs, 
                           clump = TRUE, scaled = TRUE, saveoutput = TRUE,
                           suffix = paste0("FL50_", names(subpop_v)[k]))
}
```

## Run mash
```{r}
for(k in c(1:3)){
m_out <- mash_standard_run(path = outputdir, numSNPs = numSNPs, 
                           U_hyp = U_hyp,
                           suffix = paste0("FL50_", names(subpop_v)[k]),
                           saveoutput = TRUE)
  }
```

>>>>

  #ggcov1 <- pvdiv_plot_cov(cor_phe) + ggtitle(label = paste0(phe_hyp[i], "_", names(subpop_v2)[k]))
  #ggcov <- list.prepend(ggcov, ggcov1)
  U_hyp <- list.prepend(U_hyp, cor_phe)
  names(U_hyp)[1] <- paste0(phe_hyp[i], "_", names(subpop_v2)[k])
  }
}

@@

## Set up mash inputs

### mash G&M
```{r}
i=1 # site
j=1 # phe
k=3 # subpop: G&M, M, Gulf
for(k in c(3)){
phe_suffix <- paste("FL50", subpop_v[k], 
                    sep = "*.*")
snp <- snp_attach(inputfiles$SNPfiles[k])
gwas_rds <- pvdiv_results_in_folder(path = file.path(outputdir), 
                                      pattern = paste0("GWAS_datatable*.*",
                                                       phe_suffix))
phenotypes_1 <- str_sub(gwas_rds, start = 16, end = -42) %>%
    str_replace(., "_$", "")
phenotypes <- paste0(names(subpop_v)[k], "_", phenotypes_1)
numSNPs <- round(ceiling(1200000/length(phenotypes)^2)/1000)*1000
out1phe <- pvdiv_bigsnp2mashr(path = file.path(outputdir), 
                           snp = snp, gwas_rds = gwas_rds, 
                           phenotypes = phenotypes, numSNPs = numSNPs, 
                           clump = TRUE, scaled = TRUE, saveoutput = TRUE,
                           suffix = paste0("FL50_", names(subpop_v)[k]))
}
```
Getting a lot of this: "error: chol(): decomposition failed"

## Run mash Gulf
```{r}
for(k in c(1)){
  phe_suffix <- paste("FL50", subpop_v[k], 
                      sep = "*.*")
  gwas_rds <- pvdiv_results_in_folder(path = file.path(outputdir), 
                                        pattern = paste0("GWAS_datatable*.*",
                                                         phe_suffix))
  phenotypes_1 <- str_sub(gwas_rds, start = 16, end = -42) %>%
      str_replace(., "_$", "")
  phenotypes <- paste0(names(subpop_v)[k], "_", phenotypes_1)
  numSNPs <- round(ceiling(1200000/length(phenotypes)^2)/1000)*1000
  m_out <- mash_standard_run(path = outputdir, numSNPs = numSNPs, 
                             U_hyp = U_hyp,
                             suffix = paste0("FL50_", names(subpop_v)[k]),
                             saveoutput = TRUE)
  }
```

## 2b. Midwest NB
Some univariate GWAS for the Midwest had population structure issues. Mostly for greenup - even GR50 at MO was quite bad. Drop MO for GR50. Otherwise they look ok for FL50 and GR50.

This is the error I get if I try and run mash on all eight sites with just the Midwest subpopulation as input:
Error in mashr::mash_set_data(Bhat_random, Shat_random) :
  Both Bhat and Shat are zero (or near zero) for some input data. Please check your input. If it is expected please set Shat to a positive number to avoid numerical issues; or lower the threshold to call zeros using zero_check_tol (currently set to 2.22044604925031e-16). To replace zero elements you can use `zero_Bhat_Shat_reset`.

What's happening with Midwest input?  NE & OK inputs are all zeros.
```{r}
mwdf <- readRDS(file.path(outputdir,
                          "B_hat_strong_df_19000topSNPs_FL50_Midwest.rds"))

summary(mwdf) # NE, OK random Bhats are all 0's. Shat random are all vv small.

k=2
phe_suffix <- paste("FL50", subpop_v[k], 
                    sep = "*.*")
snp <- snp_attach(inputfiles$SNPfiles[k])
gwas_rds <- pvdiv_results_in_folder(path = file.path(outputdir), 
                                      pattern = paste0("GWAS_datatable*.*",
                                                       phe_suffix))
gwas_rds <- gwas_rds[c(1:2,5:8)]

phenotypes_1 <- str_sub(gwas_rds, start = 16, end = -42) %>%
    str_replace(., "_$", "")
phenotypes <- paste0(names(subpop_v)[k], "_", phenotypes_1)
numSNPs <- round(ceiling(1200000/length(phenotypes)^2)/1000)*1000
out1phe <- pvdiv_bigsnp2mashr(path = file.path(outputdir), 
                           snp = snp, gwas_rds = gwas_rds, 
                           phenotypes = phenotypes, numSNPs = numSNPs, 
                           clump = TRUE, scaled = TRUE, saveoutput = TRUE,
                           suffix = paste0("FL50_", names(subpop_v)[k]))

U_hyp <- list()
ggcov <- list()
phe_hyp <- c("FL50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl", "crain_gr2fl", "crain_1d", "crain_2d", "crain_3d", "crain_5d", "crain_7d")

  cov_df <- phe_fl_df %>%
    ungroup() %>%
    left_join(select(sites, SITE, manu_site), by = "SITE") %>%
    filter(SUBPOP %in% subpop_v2[[k]]) %>%
    select(PLANT_ID, SUBPOP, manu_site, phe_hyp[i]) %>%
    group_by(PLANT_ID, SUBPOP, manu_site) %>%
    mutate(IND = row_number()) %>%
    pivot_wider(names_from = manu_site, values_from = phe_hyp[i]) %>%
    mutate(PLANT_ID = as.factor(PLANT_ID),
           IND = as.factor(IND)) %>%
    select(-IL) %>%
    select(PLANT_ID, SUBPOP, IND, MI, MO, SD, TX1, TX2, TX3)

  cor_phe <- cor(cov_df[,-(1:3)], use = "pairwise")
  diag(cor_phe) <- cov_df %>%
    ungroup() %>%
    summarise(MI = sqrt(var(MI, na.rm = TRUE))/mean(MI, na.rm = TRUE),
              MO = sqrt(var(MO, na.rm = TRUE))/mean(MO, na.rm = TRUE),
              SD = sqrt(var(SD, na.rm = TRUE))/mean(SD, na.rm = TRUE),
              TX1 = sqrt(var(TX1, na.rm = TRUE))/mean(TX1, na.rm = TRUE),
              TX2 = sqrt(var(TX2, na.rm = TRUE))/mean(TX2, na.rm = TRUE),
              TX3 = sqrt(var(TX3, na.rm = TRUE))/mean(TX3, na.rm = TRUE)) %>%
    as_vector()

  #ggcov1 <- pvdiv_plot_cov(cor_phe) + ggtitle(label = paste0(phe_hyp[i], "_", names(subpop_v2)[k]))
  #ggcov <- list.prepend(ggcov, ggcov1)
  U_hyp <- list.prepend(U_hyp, cor_phe)
  names(U_hyp)[1] <- paste0(phe_hyp[i], "_", names(subpop_v2)[k])
  }
}

for(k in c(2)){
  phe_suffix <- paste("FL50", subpop_v[k], 
                      sep = "*.*")
  gwas_rds <- pvdiv_results_in_folder(path = file.path(outputdir), 
                                        pattern = paste0("GWAS_datatable*.*",
                                                         phe_suffix))
  phenotypes_1 <- str_sub(gwas_rds, start = 16, end = -42) %>%
      str_replace(., "_$", "")
  phenotypes <- paste0(names(subpop_v)[k], "_", phenotypes_1)
  numSNPs <- round(ceiling(1200000/length(phenotypes)^2)/1000)*1000
  m_out <- mash_standard_run(path = outputdir, numSNPs = numSNPs, 
                             U_hyp = U_hyp,
                             suffix = paste0("FL50_", names(subpop_v)[k]),
                             saveoutput = TRUE)
  }
```

# -----------------------------
